---
---

## Storing and Re-plotting

The output of `ggplot` can be assigned to a variable, which works with `+` to
add layers.

===

```{r, handout = 0}
schl_wagp <- ggplot(person,
  aes(x = SCHL, y = WAGP, color = SEX)) +
  geom_point(
    stat = 'summary',
    fun.y = 'mean')
```

===

The plot information stored in `schl_wagp` can be used on its own, or with
additional layers.

```{r}
schl_wagp
```

===

Store additional layers by overwriting the variable (or creating a new one).

```{r, handout = 0}
schl_wagp <- schl_wagp +
  scale_color_manual(
    values = c('black', 'red'))
```
```{r}
schl_wagp
```

===

Figures are constructed in [ggplot2](){:.rlib} as layers of shapes, from the
axes on up through the `geom_*` elements. The natural file type for storing such
figures at "infinite" resolution are PDF (for print) or SVG (for online).

```{r, handout = 0}
ggsave(filename = 'schl_wagp.pdf',
  plot = schl_wagp,
  width = 4, height = 3)
```

The `plot` argument is unnecessary if the target is the most recently displayed
plot, but a little verbosity is not out-of-place here. When a raster file type
is necessary (e.g. a PNG, JPG, or TIFF) use the `dpi` argument to specify an
image resolution.
{:.notes}