---
---

## Exercises

===

### Exercise 1

Create a histogram, using a `geom_histogram` layer, of the wages earned by
females and males, with sex distinguished by the color of the bar's interior. To
silence that warning you're getting, open the help with `?geom_histogram` and
determine how to explicitly set the bin width.

===

### Exercise 2

The ACS includes data for if an individual has a science or engineering degree `SCIENGP`. Use `ggplot` to show how the mean wage earned in the U.S. varies when individuals have a science or engineering degree (coded as 1) or do not (coded as 2),
showing males and females in different colors. (Hint: Baby steps! First determine what type of plot you want to use to display categorical data for `SCIENGP`. Then
distinguish sexes by color. You can then customize your plot labels, colors, etc.)

Explore the help file for `?geom_abline` Add a horizontal line for the median wages for the data overall. 

===
### Exercise 3

The `facet_grid` layer (different from `facet_wrap`) requires an argument for
both row and column varaibles, creating a grid of panels. Create a plot with 8
facets, each displaying a single histogram of wage earned by women or men having
one of the four education attainment levels. Make the grid have 2 rows and 4
columns. *Advanced challenge*: add a second, partially transparent, histogram to
the background of each facet that provides a comparison to the whole population.
(Hint: the second histogram should not inherit the dataset from the `ggplot`
foundation.)

===

## Solutions

```{r, title = "Solution 1"}
ggplot(pums,
  aes(x = WAGP, fill = SEX)) +
  geom_histogram(binwidth = 10000)
```

```{r, title = "Solution 2"}
ggplot(na.omit(pums), aes(x=SCIENGP, y=WAGP, color=SEX)) +
  geom_boxplot()+
  scale_color_manual(values = c('blue', 'red'), 
                     labels = c('Men', 'Women')) +
  scale_x_discrete(labels = c('Yes', 'No'))+
  labs(x = "Science or Engineering Degree", 
       y = "Wages (Unadjusted USD)") +
  geom_hline(aes(yintercept= mean(WAGP)))
```

```{r, title = "Solution 3"}
ggplot(na.omit(pums),
  aes(x = WAGP)) +
  geom_histogram(bins = 20) +
  facet_grid(vars(SEX), vars(SCHL))
```

For the advanced challenge, you must supply a dataset to a second gemo_histogram
that does not have the variable specified in `facet_grid`. Note that
`facet_grid` affects the entire plot, including layers added "after faceting",
as in the solution below.

```{r, title = "Solution 3 (challenge)"}
ggplot(na.omit(pums),
  aes(x = WAGP)) +
  geom_histogram(bins = 20) +
  facet_grid(vars(SEX), vars(SCHL)) +
  geom_histogram(
    bins = 20,
    data = na.omit(pums['WAGP']),
    alpha = 0.5)
```
