---
---

## Smooth Lines

The `geom_smooth` layer can add various kinds of regression lines and
confidence intervals. The argument `method` defines the type of regression; `method = 'lm'`  specifies a linear model.

===

Here, we will observe the relationship between age (AGEP) and wages (WAGP). 

First, the scatterplot is stored as `Age_Wage`

```{r, handout = 0}
 Age_Wage <- ggplot(pums, 
       aes(x=AGEP, y=WAGP)) +
  geom_point()
```

```
Age_Wage
```

===

Next, the regression line is added using `geom_smooth` using a linear model. 

```{r, handout = 0, message = F}
Age_Wage + geom_smooth(method = 'lm')
```

===

We can also use the aesthetic mapping to explore how wages differ by sex. To do this, map the color and fill to `SEX`, using `aes` inside the `geom_smooth` layer. In `ggplot2`, `color` refers to the outline and `fill` the interior coloring.  

```{r, handout=0}
wage_gap <- Age_Wage + 
  geom_smooth(method='lm', aes(color=SEX, fill=SEX))
```

```
wage_gap
```

To learn more about adding regression lines, such as how to change the confidence interval, see the `geom_smooth` [help page] (https://ggplot2.tidyverse.org/reference/geom_smooth.html)
{:.notes}


