---
---

## Axes, labels and themes

Let's start from the histogram like the one generated in the exercise.

```{r plot_hist, title="{{ site.handouts }}"}
histo <- ggplot(data = animals_dm,
                aes(x = weight, fill = sex)) +
    geom_histogram(binwidth = 3, color = "white")
histo
```

===

We change the title and axis labels with the `labs` function. We have various functions related to the scale of each axis, i.e. the range, breaks and any transformations of the values on the axis. Here, we use `scale_x_continuous` to modify a continuous (as opposed to discrete) x-axis.

```{r plot_hist_axes, warning = FALSE, title="{{ site.handouts }}"}
histo <- histo + 
  labs(title = "Dipodomys merriami weight distribution",
       x = "Weight (g)",
       y = "Count") +
  scale_x_continuous(limits = c(20, 60),
                     breaks = c(20, 30, 40, 50, 60))
histo
```

===

For information on how to add special symbols and formatting to plot labels, see `?plotmath`.

Many plot-level options in `ggplot`, from background color to font sizes, are defined as part of *themes*. The next modification to *histo* changes the base theme of the plot to `theme_bw` (replacing the default `theme_grey`) and set a few options manually with the `theme` function. Try `?theme` for a list of available theme options.

```{r plot_hist_themes, warning = FALSE, title="{{ site.handouts }}"}
histo <- histo +
  theme_bw() +
  theme(legend.position = c(0.2, 0.5),
        plot.title = element_text(face = "bold", vjust = 2),
        axis.title.y = element_text(size = 13, vjust = 1), 
        axis.title.x = element_text(size = 13, vjust = 0))
histo
```

Note that position is relative to plot size (i.e. between 0 and 1).
