---
---

## Smooth lines

The `geom_smooth` layer adds a regression line with confidence intervals (95% CI by default). The `method = 'lm'` parameter specifies that a linear model is used for smoothing.

===

Load some data you might use for a linear model with a categorical predictor of
a continuous response.

```{r, title = '{{ site.handouts[0] }}'}
levels(animals$sex) <- c('Female', 'Male')
animals_dm <- filter(animals,
  species_id == 'DM')
```

===

With the categorical predicted mapped to an aesthetic element, the `geom_smooth`
will separately apply the `lm` method. The result hints at the importance of
the predictor.

```{r plot_lm, title = '{{ site.handouts[0] }}'}
ggplot(animals_dm,
  aes(x = year, y = weight, shape = sex)) + 
  geom_point(size = 3,
    stat = 'summary', fun.y = 'mean') +
  geom_smooth(method = 'lm')
```

===

Even better would be to distinguish everything (points and lines) by color.

```{r plot_lm_color, title='{{ site.handouts[0] }}'}
ggplot(animals_dm,
  aes(x = year, y = weight,
    shape = sex, color = sex)) + 
  geom_point(size = 3,
    stat = 'summary', fun.y = 'mean') +
  geom_smooth(method = 'lm')
```

Notice that by adding aesthetic mappings in the base aesthetic (in the `ggplot`
command), it is applied to any layer that recognizes the parameter.
{:.notes}
